Manuscript results: Forecasting influenza using ARIMA and ARIMAX

### 2025-06-20 manuscript_results.Rmd
code review up to line 2046
code review line 2900

```{r}
library("tidyr")
library("MMWRweek")
library("data.table")
library("caret")
library("purrr")
library("dplyr")
library("tseries")
library("gtools")
library("forecast")
library("scoringutils")
library("covidHubUtils")
library("parallel")
library("future")#https://cran.r-project.org/web/packages/future/vignettes/future-4-issues.html
library("listenv")
library("epitools")
library("ggplot2")
library("sf")
library("forcats")
library("ggplot2")
library("sf")
library("scales") 
library("ggplot2")
library("broom")
library("fields")
library("ggpubr")
library("patchwork")
library("ggpattern")
library("cowplot")
library("lme4")
library("ez")
```

@@

ES_1WEEK <- ggplot(W1_map) +
  geom_sf(aes(fill = AUTO_AVG_LB)) +  # Fill based on the best model
  scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 352)) +  
  ggtitle("1 week ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W1_map, aes(label = round(AUTO_AVG_LB,0)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
  labs(
    fill = "mean WIS",  
       x = "",
       y = ""
  )  # Adding a subtitle

@@

ES_2WEEK <- ggplot(W2_map) +
  geom_sf(aes(fill = AUTO_AVG_LB)) +  # Fill based on the best model
  scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 352)) +  
  ggtitle("2 weeks ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W2_map, aes(label = round(AUTO_AVG_LB,0)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
  labs(
    fill = "mean WIS",
       x = "",
       y = ""
  )  # Adding a subtitle

@@

ES_3WEEK <- ggplot(W3_map) +
  geom_sf(aes(fill = AUTO_AVG_LB)) +  # Fill based on the best model
  scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 352)) +  
  ggtitle("3 weeks ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W3_map, aes(label = round(AUTO_AVG_LB,0)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
  labs(
    fill = "mean WIS",
       x = "",
       y = ""
  )  # Adding a subtitle

@@

Loading datasets for regression analysis

@@

```{r}
library(ggpubr)

@@

Let's plot some maps with the with the mean difference (%) between the AUTO_AVG and the AUTO ARIMA models for the same state.

@@

ES_1WEEK <- ggplot(W1_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
  ggtitle("1 weeks ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W1_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
  labs(
    fill = "Mean difference",  # Label for the legend
       x = "",
       y = ""
  )  # Adding a subtitle

@@

#ggsave(ES_1WEEK,height = 4.5, width = 6.5,  filename="Fig5A.jpg")  
```
2 weeks ahead

@@

ES_2WEEK <- ggplot(W2_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
  ggtitle("2 weeks ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W2_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
  labs(
    fill = "Mean difference",  # Label for the legend
       x = "",
       y = ""
  )  # Adding a subtitle

@@

ES_3WEEK <- ggplot(W3_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
  ggtitle("3 weeks ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W3_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
  labs(
    fill =  "Mean difference",  # Label for the legend
       x = "",
       y = ""
  )  # Adding a subtitle

@@

#ggsave(ES_3WEEK, height = 4.5, width = 6.5, filename="Fig5C.jpg")  
```
4 weeks ahead

```{r}
ES_4WEEK <- ggplot(W4_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
  ggtitle("4 weeks ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W4_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
  labs(
    fill = "Mean difference",  # Label for the legend
       x = "",
       y = ""
  )  # Adding a subtitle

@@

@@

# Combine in a 2x2 grid
combined_plot <- (ES_1WEEK | ES_2WEEK) /
                 (ES_3WEEK | ES_4WEEK) + 
  plot_annotation(
    title = "Mean WIS differences compared to the baseline",
    subtitle = "Results by forecast horizon",
    theme = theme(plot.title = element_text(size = 16, face = "bold"),
                  plot.subtitle = element_text(size = 14))
  )
# Print
print(combined_plot)

@@

We will run the regression for evaluating if the best model log(WIS) which represent its performance is related to given independent variables.

@@

Computing and plotting mean WIS improvement by epidemiological weeks.

>>>>

```{r}
library("tidyr")
library("MMWRweek")
library("data.table")
library("caret")
library("purrr")
library("dplyr")
library("tseries")
library("gtools")
library("forecast")
library("scoringutils")
library("covidHubUtils")
library("parallel")
library("future")#https://cran.r-project.org/web/packages/future/vignettes/future-4-issues.html
library("listenv")
library("epitools")
library("ggplot2")
library("sf")
library("forcats")
library("ggplot2")
library("sf")
library("scales") 
library("ggplot2")
library("broom")
library("fields")
library("ggpubr")
library("patchwork")
library("ggpattern")
library("cowplot")
library("lme4")
library("ez")
library("ggpubr")
```

@@

ES_1WEEK <- ggplot(W1_map) +
  geom_sf(aes(fill = AUTO_AVG_LB)) +  # Fill based on the best model
  scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 352)) +  
  ggtitle("1 week ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W1_map, aes(label = round(AUTO_AVG_LB,0)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  
  labs(
    fill = "mean WIS",  
       x = "",
       y = ""
  )

@@

ES_2WEEK <- ggplot(W2_map) +
  geom_sf(aes(fill = AUTO_AVG_LB)) +  # Fill based on the best model
  scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 352)) +  
  ggtitle("2 weeks ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W2_map, aes(label = round(AUTO_AVG_LB,0)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
  labs(
    fill = "mean WIS",
       x = "",
       y = ""
  )

@@

ES_3WEEK <- ggplot(W3_map) +
  geom_sf(aes(fill = AUTO_AVG_LB)) +  # Fill based on the best model
  scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 352)) +  
  ggtitle("3 weeks ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W3_map, aes(label = round(AUTO_AVG_LB,0)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  
  labs(
    fill = "mean WIS",
       x = "",
       y = ""
  )

@@

Loading datasets for regression models analysis

@@

```{r}
# regression results 
print(regression_models)

@@

Let's plot some maps with the with the mean difference (%) between the AUTO_AVG_LB and the AUTO ARIMA models for the same state.

@@

ES_1WEEK <- ggplot(W1_map) +
  geom_sf(aes(fill = percentage_improvement)) +  
  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
  ggtitle("1 weeks ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W1_map, aes(label = round(percentage_improvement,0)),
               size = 3,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  
  labs(
    fill = "Mean difference", 
       x = "",
       y = ""
  )

@@

#ggsave(ES_1WEEK,height = 4.5, width = 6.5,  filename="Fig5A.jpg")  
```

@@

ES_2WEEK <- ggplot(W2_map) +
  geom_sf(aes(fill = percentage_improvement)) + 
  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
  ggtitle("2 weeks ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W2_map, aes(label = round(percentage_improvement,0)), 
               size = 3,
               color = "black",
               check_overlap = TRUE,fontface = "bold") + 
  labs(
    fill = "Mean difference",  
       x = "",
       y = ""
  )  # Adding a subtitle

@@

ES_3WEEK <- ggplot(W3_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
  ggtitle("3 weeks ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W3_map, aes(label = round(percentage_improvement,0)),  
               size = 3,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
  labs(
    fill =  "Mean difference",  # Label for the legend
       x = "",
       y = ""
  )  # Adding a subtitle

@@

#ggsave(ES_3WEEK, height = 4.5, width = 6.5, filename="Fig5C.jpg")  
```

4 weeks ahead

```{r}
ES_4WEEK <- ggplot(W4_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
  ggtitle("4 weeks ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W4_map, aes(label = round(percentage_improvement,0)), 
               size = 3,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
  labs(
    fill = "Mean difference",  # Label for the legend
       x = "",
       y = ""
  )  # Adding a subtitle

@@

Combining plot on a 2x2 grid.

@@

# Combine in a 2x2 grid
combined_plot <- (ES_1WEEK | ES_2WEEK) /
                 (ES_3WEEK | ES_4WEEK) + 
  plot_annotation(
    title = "Mean WIS differences compared to the baseline",
    subtitle = "Results by forecast horizon",
    theme = theme(plot.title = element_text(size = 16, face = "bold"),
                  plot.subtitle = element_text(size = 14))
  )

# Print
print(combined_plot)

@@

Now we will run the regression models for evaluating if the best model percentage of improvement in each state is related to given independent variables.

@@

Here we will be computing and plotting mean WIS improvement by epidemiological weeks comparing the best model and the basiline model.
